---
title: "Gaussian and Uniform geomasking"
author: "Claudio Fronterrè"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Gaussian and Uniform geomasking}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 5
)
library(geomask)
```

In this vignette we show the use of the function `geomasking` and we compare the effect of two different displacement methods on the distance of some simulated points. To illustrate this I am using a simulated dataset with 10.000 points, all located at latitude 0 and longitude 0. 

## DHS surveys geomasking method (Uniform geomasking)

This method is currenty used by the [**Demographic and Health Surveys project**](https://paperpile.com/shared/e6Csly) to protect the identity of survey respondents. GPS coordinates are displaced according to the “random direction, random distance” method following these steps:

1. Select a random direction (angle) between 0 and 360 degrees.
2. Select a random distance according to the urban and rural parameters (2km for urban and 5km for rural). 
3. Combine the results of steps 1 and 2 and assign the new coordinate to the cluster.

## Gaussian geomasking

In this case we displace the coordinates adding to them a gaussian error with mean 0 and variance $\delta^2$. 

## Simulation

We set a maximum displacement distance of 2 and we set the standard deviation of the gaussian positional error $\delta = \frac{\text{maximum distance}}{\sqrt{6}}$. This will make the two types of displacement approximately equal. These two values can be defined through the argument `delta` of the `geomasking function`. The type of geomasking applied is provided as a character string to the argument `displacement`. 

```{r, fig.align='center'}
n <- 10000
max_dist <- 2
delta <- max_dist/sqrt(6) 

# Generate the original points
true_locations <- matrix(0, nrow = n, ncol = 2)

# DHS displacement (Uniform geomasking)
unif_locs <- geomasking(locations = true_locations, displacement = "uniform", delta = max_dist)

# Gaussian geomasking
gauss_locs <- geomasking(locations = true_locations, displacement = "gaussian", delta = delta)

# Plot the displaced locations
par(mfrow = c(1, 2))
plot(unif_locs, col = rgb(50, 205, 50, alpha = 35, maxColorValue = 255), xlim = c(-2.5, 2.5), ylim = c(-2.5, 2.5), xlab = "", ylab = "", main = "Unifrom")
plot(gauss_locs, col = rgb(50, 205, 50, alpha = 35, maxColorValue = 255), xlim = c(-2.5, 2.5), ylim = c(-2.5, 2.5), xlab = "", ylab = "", main = "Gaussian")
```

```{r, fig.align='center'}
# Compare displaced distances
dist1 <- sqrt(unif_locs[, 1]^2 + unif_locs[, 2]^2)
dist2 <- sqrt(gauss_locs[, 1]^2 + gauss_locs[, 2]^2)
par(mfrow = c(1, 2))
hist(dist1, col = "lightgreen", main = "Uniform")
hist(dist2, col = "lightgreen", main = "Gaussian")

# Compare within distances of displaced points
d1 <- dist(gauss_locs)
d2 <- dist(unif_locs)
hist(d1, col = "lightgreen", xlim = c(0, 5), main = "Uniform")
hist(d2, col = "lightgreen", xlim = c(0, 5), main = "Gaussian")
```
